---
title: "Getting started with socialSim"
output:
  rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with socialSim}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(socialSim)
```

## Overview

The **socialSim** R package provides tools to simulate and analyse datasets of social interactions between individuals using hierarchical Bayesian models implemented in Stan.

This vignette demonstrates a typical workflow using three main functions:

1. `simulate_data()` – generate datasets of social interactions  
2. `run_model()` – fit a Stan model to the simulated datasets  
3. `summarise_results()` – evaluate bias and dispersion of estimated parameters

---

## 1. Simulating data

```{r simulate, eval=FALSE}
sim <- simulate_data(
  ind = 200,
  partners = 4,
  repeats = 1,
  iterations = 5,
  B_0 = 1,
  psi = 0.3,
  Valpha = 0.2,
  Vepsilon = 0.1
)
```

This creates a list of datasets representing repeated social interactions.  
You can control study design components, variance components and correlations between direct and indirect effect.

---

## 2. Fitting a model

To analyse the data, fit one of the included Stan models:

```{r model, eval=FALSE}
res <- run_model(sim, model = "Trait.stan", iter = 2000, cores = 4)
```

Importantly, you will need **rstan** installed for this step.

---

## 3. Summarising results

Once the models are fitted, summarise bias and dispersion across simulations:

```{r summarise, eval=FALSE}
summary <- summarise_results(res)
print(summary)
```

This function extracts model estimates and computes metrics such as mean absolute deviation (MADm) across replicates.

---

## Example output (simulated workflow)

Here’s a minimal example with few iterations for a fast runtime:

```{r example, eval=FALSE}
sim <- simulate_data(ind = 50, partners = 2, iterations = 4, Valpha = 0.2, Vepsilon = 0.1)
res <- run_model(sim, model = "Trait.stan", iter = 500, cores = 4)
summary <- summarise_results(res)
print(summary)
```

---

## Conclusion

The **socialSim** package helps researchers design, simulate, and evaluate models of social phenotypes and indirect genetic effects.

```{r}
?simulate_data
?run_model
?summarise_results
```

and visit the [GitHub page](https://github.com/RoriWijnhorst/socialSim) for the latest updates.
